---
title: 'Homework for class N01: Diagnostic test example'
output:
  html_document:
    df_print: paged
---

# Homework for class N01.

```{r include=FALSE}
# Removing any previous answers:
previousAnswerVariables = grep(pattern=('Answer[[:digit:]]+'), ls(), v=T) 
if(length(previousAnswerVariables) > 0) rm(list=previousAnswerVariables)
```

#Table 1: diagnosticModelFamily


```{r}
Table1_data = diagnosticModelFamily = 
  rbind(c(0.95, 0.03, 0.02), 
        c(0.03, 0.95, 0.02)) 
parameterSpace = thetaValues = c("healthy","sick")
sampleSpace = Xvalues = c("negative","positive","indeterminate")
dimnames(diagnosticModelFamily) = dimnames(Table1_data) = 
  list(theta=parameterSpace, X=sampleSpace)
print(diagnosticModelFamily)
```

### QUESTION 1a:  Which variable, X or theta, are the table entries conditional on?
(Replace NULL with your answer, in this chunk.)

```{r}
Answer1a = 'theta'
```
Here are the row sums (could be informative!):

```{r}
print(rowSums(Table1_data))
print(colSums(Table1_data))

```


### QUESTION 1b: What are the sensitivity and specificity of the test, if "indeterminate" is combined with "positive"? 
Don't just calculate them.  Write a function to do it.

```{r}
SensitivitySpecificity = function(
  diagnosticModelFamily = Table1_data,
  which.positive=c("positive","indeterminate"),
  which.negative=c("negative")
) {
  diagnosticModelFamilyCollapsed = cbind(
    rowSums(diagnosticModelFamily[ , which.negative, drop=F]), 
    rowSums(diagnosticModelFamily[ , which.positive, drop=F])
  ) 
  dimnames(diagnosticModelFamilyCollapsed)[[2]] = 
    c('negative','positive')
  sens = diagnosticModelFamilyCollapsed[2,2]
  sens = diagnosticModelFamilyCollapsed['sick', 'positive']
  spec = diagnosticModelFamilyCollapsed['healthy', 'negative']
  return(c(sensitivity=sens, specificity=spec))  
}

Answer1b = SensitivitySpecificity()
Answer1b
```


### QUESTION 1c: What are the sensitivity and specificity of the test, if "indeterminate" is combined with "negative"? 
Don't just calculate them.  Write a simple function to do it.

```{r}
Answer1c = SensitivitySpecificity(diagnosticModelFamily = Table1_data, which.positive = "positive", which.negative = c("negative" , "indeterminate"))
```

### QUESTION 1d: What are the likelihood ratios for the three possible obervations (values of X)?

```{r}
Answer1d = Table1_data[2, ]/Table1_data[1, ]
Answer1d
```


#Table 2:  Joint probability table

To get the joint probability TABLE, we multiply the Table 1 bottom row by the prevalence, and the top row by 1 - prevalence.

```{r}
prior = c(healthy=0.9, sick=0.1)
Table_2_data = diagnosticJointTable = 
  matrix(prior, nrow=2, ncol=3) *
    diagnosticModelFamily
print(diagnosticJointTable)
```

QUESTION 2a:  Write code that verifies that all the probabilities in diagnosticJointTable add to 1.

```{r}
Answer2a = sum(Table_2_data = diagnosticJointTable)
```


QUESTION 2b:  Write a function "joint_probability" that takes as its arguments an outcome X  and a state-of-nature, and returns the probability of that outcome.

Example of use: joint_probability("negative", "healthy")

HINT:  It should be just ONE LINE, very simple. Use indexing. 
Type help('[') at the console to learn about indexing.



```{r}
Answer2b = joint_probability=function(X,theta){
  diagnosticJointTable[theta, X]
}
joint_probability("negative", "healthy")
```



# Report of Answers:

```{r}
AnswerVariables = grep(pattern=('Answer[[:digit:]]+'), ls(), v=T) 
print(AnswerVariables)
Answers = sapply(AnswerVariables, get)
print(Answers)
```

